### 2023-06-02 paper/localpsf.tex
in the middle of editing paper

Once the batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed $\impulseresponse_{x_i}$ (Section~\ref{sec:approximate_kernel_entries}). The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section~\ref{sec:local_mean_displacement_invariance}. Specifically, we approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a small number of sample points $x_i$ near $x$. The weights are determined by radial basis function interpolation.

@@

We start with a finite set of candidate points $\candidatepts$ and build $\pointbatch_k$ incrementally with points selected from $\candidatepts$. For simplicity of explanation, here $\pointbatch_k$ and $\candidatepts$ are mutable sets that we add points to and remove points from. First we initialize $\pointbatch_k$ as an empty set. Then we select the candidate point $\candidatepoint_i \in \candidatepts$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$. Candidate points for the first batch $S_1$ are chosen randomly from $\candidatepts$.
Once $\candidatepoint_i$ is selected, we remove $\candidatepoint_i$ from $\candidatepts$. Then we perform the following checks:
\begin{enumerate}
	\item We check whether $\candidatepoint_i$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ for all $\candidatepoint_j \in \pointbatch_k$.
	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon_V \spatialvol_\text{max}$. Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon_V$ is a small threshold parameter (we use $\epsilon_V=10^{-5}$).
	\item We make sure that all eigenvalues of $\spatialcov(\candidatepoint_i)$ are positive, and the aspect ratio of $E_{\candidatepoint_i}$ (ratio of the largest eigenvalue of $\spatialcov(\candidatepoint_i)$ to the smallest) is bounded by a constant $1/\epsilon_\Sigma$ (we use $1/\epsilon_\Sigma=20$). Negative integral kernel entries due to discretization error can cause $\spatialcov(\candidatepoint_i)$ to be indefinite or highly ill-conditioned. We reject these points because it is likely that the set $E_{\candidatepoint_i}$ poorly reflects the support of $\phi_{\candidatepoint_i}$.
\end{enumerate}
If $\candidatepoint_i$ passes these checks 
%(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ is not too small, and the )
then we add $\candidatepoint_i$ to $\pointbatch_k$. Otherwise we discard $\candidatepoint_i$. This process repeats until there are no more points in $\candidatepts$.  
%This is detailed in Algorithm~\ref{alg:point_choice}. 
%We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the fast ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}. 
We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$. For each batch, $\candidatepts$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previous batches.

@@

		\includegraphics[scale=0.13]{mstar_0.25noise_cropped.png}

@@

		\includegraphics[scale=0.13]{mstar_0.05noise_cropped.png}

@@

	\includegraphics[scale=0.13]{mstar_0.01noise_cropped.png}

@@

\appendix

@@

\section{Ellipsoid intersection test}
\label{sec:fast_ellipsoid_intersection_test}
The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
\begin{align*}
	E_i :=& \{x : (x - \spatialmean_i)^T \spatialcov_i^{-1} (x - \spatialmean_i) \le \tau^2\}
\end{align*}
for $i\in\{1,2\}$, where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite. Let $K$ be the following one dimensional convex function:
\begin{equation*}
	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2).
\end{equation*}
In~\cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
% (see Figure~\ref{fig:ellipsoid_intersection_test}).

%\begin{figure}
%	\begin{subfigure}{0.37\textwidth}
%		\centering
%		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
%	\end{subfigure}
%	\begin{subfigure}{0.30\textwidth}
%		\centering
%		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
%	\end{subfigure}
%	\begin{subfigure}{0.30\textwidth}
%		\centering
%		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
%	\end{subfigure}
%	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
%	}
%	\label{fig:ellipsoid_intersection_test}
%\end{figure}

The function $K(s)$ may be evaluated quickly for many $s$ by computing the solution to the generalized eigenvalue problem
$
	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
$
where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors, and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. 
%The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix. Using this diagonalization, and 
After some algebraic manipulations, we have
\begin{equation}
	\label{eq:Ks_generalized}
	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$. To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form~\eqref{eq:Ks_generalized} on $(0,1)$ using Brent's method~\cite{Brent71}. If $K(s^*) <0$ at the minimizer, $s^*$, then $E_1 \cap E_2 = \{\}$. Otherwise $E_1 \cap E_2 \neq \{\}$.
%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm~\ref{alg:ellipsoid_intersection_test}.

%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	\SetKwProg{Fn}{Function}{}{}
%	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
%		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
%	}
%	
%	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
%	
%	
%	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
%		
%	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
%		
%	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
%		
%	\If{$K^* < 0$}{
%			
%		$\text{ellipsoids\_intersect} \gets \text{False}$
%			
%	}
%	\Else{
%			
%		$\text{ellipsoids\_intersect} \gets \text{True}$
%			
%	}
%	\caption{Determining whether two ellipsoids intersect}
%	\label{alg:ellipsoid_intersection_test}
%\end{algorithm2e}

%\subsection{Ellipsoid bounding box test}

This ellipsoid intersection test is only performed if the coordinate axis aligned bounding boxes for the ellipsoids intersect. The smallest coordinate axis aligned box that contains $E_i$ is the box
$
	\prod_{k=1}^\gdim \left[ \spatialmean_i^{k} -\tau\sqrt{\Sigma_i^{kk}}, \spatialmean_i^{k} +\tau\sqrt{\Sigma_i^{kk}} \right],
$
where $\spatialmean_i^{k}$ is the $k$th entry of $\spatialmean_i$, and $\Sigma_i^{kk}$ is the $k$th diagonal entry of $\Sigma_i$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
$\prod_{k=1}^\gdim \left[a_{-}^k, a_{+}^k\right]$ and $\prod_{k=1}^\gdim \left[b_{-}^k, b_{+}^k\right]$
intersect if and only if $a_{-}^k \le b_{+}^k$ and $b_{-}^k \le a_{+}^k$ for $k=1,\dots,d$.

>>>>

Once the batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed $\impulseresponse_{x_i}$ (Section~\ref{sec:approximate_kernel_entries}). The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section~\ref{sec:local_mean_displacement_invariance}. Specifically, we approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a small number of sample points $x_i$ near $x$. The weights are determined by radial basis function (RBF) interpolation.

@@

We start with a finite set of candidate points $\candidatepts$ and build $\pointbatch_k$ incrementally with points selected from $\candidatepts$. For simplicity of explanation, here $\pointbatch_k$ and $\candidatepts$ are mutable sets that we add points to and remove points from. First we initialize $\pointbatch_k$ as an empty set. Then we select the candidate point $\candidatepoint_i \in \candidatepts$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$. Candidate points for the first batch $S_1$ are chosen randomly from $\candidatepts$.
Once $\candidatepoint_i$ is selected, we remove $\candidatepoint_i$ from $\candidatepts$. Then we perform the following checks:
\begin{enumerate}
	\item We check whether $\candidatepoint_i$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ for all $\candidatepoint_j \in \pointbatch_k$.
	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon_V \spatialvol_\text{max}$. Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon_V$ is a small threshold parameter (we use $\epsilon_V=10^{-5}$).
	\item We make sure that all eigenvalues of $\spatialcov(\candidatepoint_i)$ are positive, and the aspect ratio of $E_{\candidatepoint_i}$ (ratio of the largest eigenvalue of $\spatialcov(\candidatepoint_i)$ to the smallest) is bounded by a constant $1/\epsilon_\Sigma$ (we use $1/\epsilon_\Sigma=20$). Negative integral kernel entries due to discretization error can cause $\spatialcov(\candidatepoint_i)$ to be indefinite or highly ill-conditioned. We reject these points because it is likely that the set $E_{\candidatepoint_i}$ poorly reflects the support of $\phi_{\candidatepoint_i}$.
\end{enumerate}
If $\candidatepoint_i$ passes these checks 
(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ and $\spatialcov(\candidatepoint_i)$ are acceptable)
then we add $\candidatepoint_i$ to $\pointbatch_k$. Otherwise we discard $\candidatepoint_i$. This process repeats until there are no more points in $\candidatepts$.  
%This is detailed in Algorithm~\ref{alg:point_choice}. 
We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the fast ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
%described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}. 
We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$. For each batch, $\candidatepts$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previous batches.

@@

		\includegraphics[scale=0.20]{mstar_0.25noise_cropped.png}

@@

		\includegraphics[scale=0.20]{mstar_0.05noise_cropped.png}

@@

	\includegraphics[scale=0.20]{mstar_0.01noise_cropped.png}

@@

%\appendix

@@

%\section{Ellipsoid intersection test}
%\label{sec:fast_ellipsoid_intersection_test}
%The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
%	E_i :=& \{x : (x - \spatialmean_i)^T \spatialcov_i^{-1} (x - \spatialmean_i) \le \tau^2\}
%for $i\in\{1,2\}$, where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite. Let $K$ be the following one dimensional convex function:
%	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2).
%In~\cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
%% (see Figure~\ref{fig:ellipsoid_intersection_test}). 
%%	\begin{subfigure}{0.37\textwidth}
%%		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
%%	\end{subfigure}
%%	\begin{subfigure}{0.30\textwidth}
%%		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
%%	\end{subfigure}
%%	\begin{subfigure}{0.30\textwidth}
%%		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
%%	\end{subfigure}
%%	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
%%	}
%%	\label{fig:ellipsoid_intersection_test}
%The function $K(s)$ may be evaluated quickly for many $s$ by computing the solution to the generalized eigenvalue problem
%	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
%where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors, and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. 
%%The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix. Using this diagonalization, and 
%After some algebraic manipulations, we have
%	\label{eq:Ks_generalized}
%	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
%where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$. To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form~\eqref{eq:Ks_generalized} on $(0,1)$ using Brent's method~\cite{Brent71}. If $K(s^*) <0$ at the minimizer, $s^*$, then $E_1 \cap E_2 = \{\}$. Otherwise $E_1 \cap E_2 \neq \{\}$.
%%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm~\ref{alg:ellipsoid_intersection_test}.
%%\begin{algorithm2e}
%%	\SetAlgoNoLine
%%	\SetKwInOut{Input}{Input}
%%	\SetKwInOut{Output}{Output}
%%	\SetKwProg{Fn}{Function}{}{}
%%	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
%%		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
%%	}
%%	
%%	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
%%	
%%	
%%	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
%%		
%%	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
%%		
%%	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
%%		
%%	\If{$K^* < 0$}{
%%			
%%		$\text{ellipsoids\_intersect} \gets \text{False}$
%%			
%%	}
%%	\Else{
%%			
%%		$\text{ellipsoids\_intersect} \gets \text{True}$
%%			
%%	}
%%	\caption{Determining whether two ellipsoids intersect}
%%	\label{alg:ellipsoid_intersection_test}
%%\end{algorithm2e}
%%\subsection{Ellipsoid bounding box test}
%This ellipsoid intersection test is only performed if the coordinate axis aligned bounding boxes for the ellipsoids intersect. The smallest coordinate axis aligned box that contains $E_i$ is the box
%	\prod_{k=1}^\gdim \left[ \spatialmean_i^{k} -\tau\sqrt{\Sigma_i^{kk}}, \spatialmean_i^{k} +\tau\sqrt{\Sigma_i^{kk}} \right],
%where $\spatialmean_i^{k}$ is the $k$th entry of $\spatialmean_i$, and $\Sigma_i^{kk}$ is the $k$th diagonal entry of $\Sigma_i$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
%$\prod_{k=1}^\gdim \left[a_{-}^k, a_{+}^k\right]$ and $\prod_{k=1}^\gdim \left[b_{-}^k, b_{+}^k\right]$
%intersect if and only if $a_{-}^k \le b_{+}^k$ and $b_{-}^k \le a_{+}^k$ for $k=1,\dots,d$.
